DOM bridges and barriers: systematic review and survey

### 2025-09-24 reports/draft_results.qmd
time series and heat map

## Result 2: Frequencies according to review and survey

@@

@@

### 3.1: Conservation problems DOM has been used to address

>>>>

## Result 2: Codes by review and survey

### Result 2.1: Frequencies according to review and survey

@@

### Result 2.2: Time series of top theme from literature, co-management

Grey is the number of DOM publications per year, while blue is the number of times co-management and communication was referenced. Whanted to see if this code was referenced more over time, but seems that it just tracks the number of DOM publications. Likely won't include.

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 7
#| fig-height: 5
#| fig-align: center

code_ts <- codes_vars %>% 
  mutate(year = str_extract_all(Doc, "\\d+\\.?\\d*"), 
         year = as.Date(as.character(year), format = "%Y"), 
         year = year(year)) %>%
  select(c(Doc, year, Co.management.and.communication)) %>%
  group_by(year) %>%
  summarise(n_appear = sum(Co.management.and.communication), 
            n_year = n())

colnames(code_ts) <- c("Year", "com", "pubs")

ggplot(code_ts, aes(as.numeric(Year))) + 
  geom_path(aes(y = com), linewidth = 1.5, color = "dodgerblue4") + 
  geom_point(aes(y= com), size = 5, fill = "dodgerblue4", color = "black", shape = 21) + 
  geom_path(aes(y = pubs), linewidth = 1.5, color = "grey60") + 
  geom_point(aes(y = pubs), size = 5, fill = "grey60", color = "black", shape = 21) +
  xlab("") +
  ylab("Count (#)") +
  theme_minimal() + 
  theme(axis.text = element_text(size = 14, color = "black"), 
        axis.title = element_text(size = 16, color = "black")) 
```

### Result 2.3: Heat maps of theme co-occurence

This heatmap represents number of times two codes co-occurered together in the same article, or "theme co-occurence". I have filtered totals greater than or equal to 11, which was the 90% quantile of the co-occurrence total data set. Blue text represents data bridges and barriers, grey text represents ecological bridges and barriers, and orange text represents social bridges and barriers.

```{r}
#| echo: false
#| message: false
#| warning: false
#| fig-width: 5
#| fig-height: 5
#| fig-align: center
#| include: FALSE

code_rel_melt <- reshape::melt(code_rel)
code_rel_melt$variable <- gsub(".", " ", code_rel_melt$variable, fixed = TRUE)

code_rel_melt <- code_rel_melt %>%
  separate(Code.System, c("key", "code1"), "> ", convert = TRUE)

code_rel_melt$key <- str_trim(code_rel_melt$key)

colnames(code_rel_melt) <- c("brba_code1", "code1", "code2", "total")

#add brba column to second code col
code_rel_melt2 <- code_rel_melt
code_rel_melt2$brba_code2 <- NA

code_rel_melt2$code2[code_rel_melt2$code2 == "Fine scale management appropriate for target species"] <- "Fine-scale management appropriate for target species"
code_rel_melt2$code2[code_rel_melt2$code2 == "Pre existing regulation and management institutions"] <- "Pre-existing regulation and management institutions"
code_rel_melt2$code2[code_rel_melt2$code2 == "Co management and communication"] <- "Co-management and communication"

for(i in 1:nrow(code_rel_melt2)){
  code_temp = code_rel_melt2$code2[i]

  brba_temp = code_rel_melt2 %>% filter(code1 == code_temp & code2 == code_temp)
  brba_code_temp = brba_temp$brba_code1[1]

  code_rel_melt2$brba_code2[i] <- brba_code_temp
}

code_rel_melt2$brba_code2[code_rel_melt2$code2 == "Target species is a prey specialist"] <- "Ecological bridge"

group_colors = c("Data Bridge" = "dodgerblue4", "Data Barrier" = "dodgerblue4", "Ecological Bridge" = "honeydew4", "Ecological Barrier" = "honeydew4", "Social Bridge" = "darkorange4", "Social Barrier" = "darkorange4")

code_rel_melt_h <- code_rel_melt2 %>%
  filter(total >= 11) %>%
  mutate(bor_box = ifelse(total >= 11, "TRUE", "FALSE"), 
        bor_box = as.factor(bor_box), 
        code1 = as.factor(code1), 
        code2 = as.factor(code2))

code_rel_melt_h$bor_box <- factor(code_rel_melt_h$bor_box, levels = c("TRUE", "FALSE"))

code_rel_melt_h <- code_rel_melt_h %>%
  mutate(code1 = reorder(code1, -total, FUN = mean),
         code2 = reorder(code2, total, FUN = mean))

unique_x_groups <- code_rel_melt_h %>%
distinct(code1, brba_code1) %>%
arrange(code1)

unique_y_groups <- code_rel_melt_h %>%
distinct(code2, brba_code2) %>%
arrange(code2)

x_colors <- setNames(group_colors[unique_x_groups$brba_code1], unique_x_groups$code1)
y_colors <- setNames(group_colors[unique_y_groups$brba_code2], unique_y_groups$code2)

ggplot(code_rel_melt_h, aes(x = code1, y = reorder(code2, total))) +
geom_tile(aes(fill = total, colour = bor_box)) + 
scale_fill_gradient2(high = "red")+
ylab("")+
xlab("")+
labs(fill = "Number of co-occurences")+
theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.3, size = 14, face = "bold", color = x_colors[levels(code_rel_melt_h$code1)]), 
      axis.text.y = element_text(size = 14, face = "bold", color = y_colors[levels(code_rel_melt_h$code2)]), 
      panel.background = element_blank())+
geom_text(aes(label = total))+
scale_color_manual(values = c("black", "#00000000"))+
guides(color = "none")

#ggsave("figures/ms_figs/heat_tile.png", width = 10, height = 8, units = c("in"))

![](images/heat_tile_ai.png){fig-align="center" width="10in", height = "8in"}

@@

### 3.1: Management and conservation problems DOM has been used to address
